---
title: "Practical Machine Learning Course Project"
author: "Marsela Yulita"
date: "20 September 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE )
```

## Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

#### Load the data
```{r load}
library(caret)
library(randomForest)
library(rattle)


#Download the data
if(!file.exists("pml-training.csv")){
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv")}

if(!file.exists("pml-testing.csv")){
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")}

#Read the data
trainingData<- read.csv("pml-training.csv", sep=",", header=TRUE, na.strings = c("NA","",'#DIV/0!'))
testingData<- read.csv("pml-testing.csv", sep=",", header=TRUE, na.strings = c("NA","",'#DIV/0!'))

dim(trainingData)
dim(testingData)
```

#### Preprocessing data
Remove the missing values and zero variance variable.

```{r preproses}
#Remove fields with NA
trainingData <- trainingData[,(colSums(is.na(trainingData)) == 0)]
testingData <- testingData[,(colSums(is.na(testingData)) == 0)]

dim(trainingData)
dim(testingData)

#Preprocess step
numericalcol <- which(lapply(trainingData, class) %in% "numeric")

preprocessModel <-preProcess(trainingData[,numericalcol],method=c('knnImpute', 'center', 'scale'))
pre_trainingData <- predict(preprocessModel, trainingData[,numericalcol])
pre_trainingData$classe <- trainingData$classe

pre_testingData <-predict(preprocessModel,testingData[,numericalcol])

#Remove NearZeroVariance variables
nzv <- nearZeroVar(pre_trainingData,saveMetrics=TRUE)
pre_trainingData <- pre_trainingData[,nzv$nzv==FALSE]

nzv <- nearZeroVar(pre_testingData,saveMetrics=TRUE)
pre_testingData <- pre_testingData[,nzv$nzv==FALSE]
```

#### Cross validation
I will create partition in data set to make a training data set and test data set.

```{r crossval}
#create data partition
set.seed(2017)
dtTrain<- createDataPartition(pre_trainingData$classe, p=0.6, list=FALSE)
training<- pre_trainingData[dtTrain, ]
validation <- pre_trainingData[-dtTrain, ]
dim(training) ; dim(validation)
```

#### Train Data
##### Decision Tree (rpart)
Then, I will used decision tree to build a predictive model.

```{r trainrpart}
#train model with rpart
rpartmodel <- train(classe~., method="rpart", data=training)
rpartmodel$finalModel
fancyRpartPlot(rpartmodel$finalModel,cex=.5,under.cex=1,shadow.offset=0)

#predict with validation data
rpartpredict <- predict(rpartmodel,validation)
rpartmatrix <- confusionMatrix(validation$classe,rpartpredict)
rpartmatrix
```

Unfortunately, we got the accuracy only 48.5%.

##### Random forest
Let's try the random-forest technique to generate a predictive model.

```{r trainrf}
#train model with random forest
modelFit <- train(classe ~., method="rf", data=training, trControl=trainControl(method='cv'), number=5, allowParallel=TRUE, importance=TRUE )
modelFit
plot(modelFit)

#predict with validation data
predictvalid <- predict(modelFit, validation)
validmatrix <- confusionMatrix(validation$classe, predictvalid)
validmatrix
```
We got 99.7% accuracy for random forest training model. So this model was a better model than decision tree.

#### Evaluate the model (in and out of sample error)
```{r eva}
#In training set
insample <- predict(modelFit,training)
confusionMatrix(training$classe,insample)

#out sample
confusionMatrix(validation$classe, predictvalid)
```

The accuracy of the prediction is 99.07%. Hence, the in and out of sample error is 100% - 99.07% = 0.93%.

#### Using the model with the test data
```{r test}
pred_final <- predict(modelFit, pre_testingData)
pred_final
```

Here are the results, we will use them for the submission of this course project in the coursera platform.
